---
title: "Data manipulation2"
author: "Xicheng Xie"
date: "2022-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r }
library(tidyverse)

options(tibble.print_min = 5)
```

## Data import

```{r pressure, echo=FALSE}
pulse_df<-haven::read_sas("data_import_examples/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()

```

## use of `pivot_longer`
```{r}
pulse_tidy_data<-
  pivot_longer(
    pulse_df,
    bdi_score_bl:bdi_score_12m,
    names_to = 'visit',
    values_to = 'bdi'
  )
```

## drop the prefilx
```{r}
pulse_tidy_data<-
  pivot_longer(
    pulse_df,
    bdi_score_bl:bdi_score_12m,
    names_to = 'visit',
    names_prefix = 'bdi_score_',
    values_to = 'bdi'
  )
```

## replace, bi with 00 month, use replace in the `mutate` and `factor`
```{r}
pulse_df = 
  haven::read_sas("data_import_examples/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  relocate(visit) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"), #??
    visit = factor(visit)) %>%
  arrange(id, visit)
```

##**Learning Assessment**: 
In the litters data, the variables gd0_weight and gd18_weight give the weight of the mother mouse on gestational days 0 and 18. Write a data cleaning chain that retains only litter_number and these columns; produces new variables gd and weight; and makes gd a numeric variable taking values 0 and 18 (for the last part, you might want to use recode …). Is this version “tidy”?

```{r}
litter_tidy_data<-litters_data %>% 
  janitor::clean_names() %>% 
  select(litter_number, ends_with("weight")) %>% 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = 'gd',
    values_to = 'weight'
  ) %>% 
  mutate(gd=recode(gd,'gd0_weight'=0,'gd18_weight'=18)) # !!!
```

## `pivot_wider`
```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean")
```
## Binding rows
```{r}
fellowship_ring<-readxl::read_excel('data_import_examples/LotR_Words.xlsx',range = "B3:D6") %>% 
  mutate(movie="fellowship_ring")

two_towers = 
  readxl::read_excel("./data_import_examples/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data_import_examples/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
lotr_tidy<-bind_rows(fellowship_ring,two_towers,return_king) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    female:male,
    names_to = 'gender',
    values_to = 'words') %>% 
    mutate(race = str_to_lower(race)) %>% 
  select(movie, everything()) 
```


